---
title: "P8105 Homework 1"
author: "Stephanie Chuquipul Calderon"
date: "2026-09-26"
output: github_document
---

# Problem 1 

```{r}
data("penguins", package = "palmerpenguins")
```
This dataset contains the following penguin species: Adelie, Gentoo, and Chinstrap. These penguins are from the islands of Torgersen, Biscoe, and Dream. The dataset also contains the bill length and depth in millimeters, flipper length in millimeters, body mass in grams, sex and year. 
The penguin dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. 
The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 1)` mm.

```{r}
penguin_plot =
ggplot2::ggplot(
  data = penguins,
  mapping = ggplot2::aes(
    x = bill_length_mm,
    y = flipper_length_mm,
    color = species
  )
) +
  ggplot2::geom_point()

penguin_plot
ggplot2::ggsave (
  filename = "penguin_scatterplot.png",
  plot = penguin_plot
)
```

# Problem 2

```{r}
set.seed(123)

normal_sample = rnorm(10)
greater_than_zero = normal_sample > 0
greater_than_zero
letter_labels = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J")
letter_labels
size_group = factor(c(
  "small", "medium", "large", "small", "medium",
  "large", "small", "medium", "large", "small"
))
size_group
```

```{r}
homework1_df = data.frame(
  normal_sample = normal_sample,
  greater_than_zero = greater_than_zero,
  letter_labels = letter_labels,
  size_group = size_group
)

homework1_df
mean(homework1_df$normal_sample)
mean(homework1_df$greater_than_zero)
mean(homework1_df$letter_labels)
mean(homework1_df$size_group)
```


```{r convert-types, results='hide', warning=FALSE}
as.numeric(homework1_df$greater_than_zero)
as.numeric(homework1_df$letter_labels)
as.numeric(homework1_df$size_group)
```

The mean of the numeric sample is about 0.075. The mean of the logical column is 0.5 because R counts TRUE as 1 and FALSE as 0; five of the ten values are TRUE. The means of the character and factor columns return NA because their values are categories rather than numbers. With as.numeric(), the logical values become 1s and 0s, while the letters become NA because they cannot be read as numbers. The factor becomes numeric codes for its three levels, but those codes are not measurements, so averaging them would not be meaningful.









